---
title: "Predicting Activity Classe"
output: html_document
---

## Introduction

The goal of this project is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.

## Data Description
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

## Data Preparation

```{r libraries and datasets, echo=T}
# Load libraries
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)

set.seed(124)

# Set the URL for the download
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# Download the datasets
training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))

# Create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
# Check dimensions
dim(TrainSet)
dim(TestSet)

# Remove variables with Nearly Zero Variance
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
# Check dimensions
dim(TrainSet)
dim(TestSet)

# Remove variables that are mostly NA
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
# Check dimensions
dim(TrainSet)
dim(TestSet)

# Remove identification only variables (columns 1 to 5)
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
# Check dimensions
dim(TrainSet)
dim(TestSet)
```
## Model option 1: Random Forest

```{r random forest, echo=T}
# Fit the model
controlRF <- trainControl(method="cv", number=3, verboseIter=F)
RF_mod <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
RF_mod$finalModel

# Prediction on test dataset
predict_RF <- predict(RF_mod, newdata=TestSet)
# Confusion Matrix
conf_RF <- confusionMatrix(table(predict_RF, TestSet$classe))
conf_RF

```

## Model Option #2: GBM
```{r gbm, echo=T}
# Fit the model
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
GBM_mod  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = controlGBM, verbose = F)
GBM_mod$finalModel

# prediction on Test dataset
predict_GBM <- predict(GBM_mod, newdata=TestSet)
conf_GBM <- confusionMatrix(table(predict_GBM, TestSet$classe))
conf_GBM

```

## Conclusion
The Random Forest model has slightly higher accuracy (.9978) than GBM (.9866). Hence, Random Forest model will be used for the final prediction.
                
```{r predict, echo=T}
predictTEST <- predict(RF_mod, newdata=testing)
predictTEST
```
